---
editor_options: 
  chunk_output_type: console
---

# Eye size analyses 

In the first section of the script, we will compare the eye size estimates obtained using photos from Macaulay Library with Ausprey & Ritland 2024's eye morphology dataset from preserved museum specimens. This is followed by the second section of the script where we will focus on obtaining mean eye size estimates from citizen-science photographs for all species in our study.

## Install required libraries  

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(sf)
library(raster)
library(ggpubr)
```

## Loading the data

```{r}
species_traits <- read.csv ('data/species-trait-dat.csv') 
eye_size_museum <- read.csv ('data/eye-size-museum.csv') #file with museum-derived eye size measurements
eye_size_ebird_AR <- read.csv ('data/eye-size-ebird-A&R.csv') #file with Macaulay photo-derived eye size estimated, for species for which we have comparable measurements from museum specimens
eye_size_ebird_not_in_AR <- read.csv ('data/eye-size-ebird-not-in-A&R.csv') #file with Macaulay photo-derived eye size estimates, for species for which we don't have comparable measurements from museum specimens
annotations <- read.csv ('results/data.csv') 
```

## Updating the scientific name in the species_traits data

```{r}
species_traits <- species_traits %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) #updating the scientific name of yellow-browed bulbul
```

## Cleaning the eye size data obtained from eBird photos

```{r}
eye_size_ebird_AR <- eye_size_ebird_AR %>%
  rename (scientific_name = Species) %>% #renaming for easier analyses
  dplyr::select (scientific_name, Eye_Size_in_mm_averaged, Error_of_bill_ratio) %>% #retaining only select columns
 group_by (scientific_name) %>%
  slice_min (order_by = abs(Error_of_bill_ratio), n = 10) %>% #only selecting those estimates which have the least error ratios of all estimates
  summarize (mean_eye_size = mean(Eye_Size_in_mm_averaged)) %>% #obtaining the mean eye size measurements for our species
  left_join (., species_traits, by = 'scientific_name') %>%
  dplyr::select (scientific_name, mean_eye_size, common_name, mass)  %>%
  filter (common_name %in% annotations$common_name) %>%
  ungroup ()

eye_size_ebird_not_in_AR <- eye_size_ebird_not_in_AR %>%
  na.omit () %>%
  filter (!common_name %in% c("Red-wattled Lapwing", "Spotted Dove", "Malabar Parakeet", "Vernal Hanging-Parrot", "Plum-headed Parakeet", "Malabar Gray Hornbill")) %>% #removing 6 species for which either the error rates are too high, or eye sizes are not possible to obtain using photographs
  rename (scientific_name = species) %>% 
  dplyr::select (scientific_name, Eye_Size_in_mm_averaged, Error_of_bill_ratio) %>% 
  group_by (scientific_name) %>%
  mutate (Error_of_bill_ratio = as.numeric (Error_of_bill_ratio), Eye_Size_in_mm_averaged = as.numeric (Eye_Size_in_mm_averaged)) %>% #converting into numeric class
  slice_min (order_by = abs(Error_of_bill_ratio), n = 10) %>% 
  summarize (mean_eye_size = mean(Eye_Size_in_mm_averaged)) %>% 
  left_join (., species_traits, by = 'scientific_name') %>%
  dplyr::select (scientific_name, mean_eye_size, common_name, mass) %>% 
  ungroup ()
```

## Obtaining a list of species that are present and absent in the eye size measurements dataset- only for reference

```{r}
#species present in annotations data for which Ausprey and Ritland HAVE eye size data from museum specimens
species_in_data <- data.frame (scientific_name = intersect (unique (annotations$scientific_name), eye_size_museum$species_jetz)) #yields 38 species

setdiff (species_in_data$scientific_name, eye_size_ebird_AR$scientific_name) #yields 2 species- Great Hornbill and Jungle Nightjar for which obtaining eye size estimates from photos was not possible. So, we have eye size measurements from museum and eye size estimates from photos for 36 species. 

#species present in annotations data for which Ausprey and Ritland DO NOT HAVE eye size data from museum specimens
species_not_in_data <- data.frame (scientific_name = setdiff (unique (annotations$scientific_name), eye_size_museum$species_jetz)) #yields 45 species

setdiff (species_not_in_data$scientific_name, eye_size_ebird_not_in_AR$scientific_name) #yields 6 species- Red-wattled Lapwing, Malabar Grey Hornbill, Malabar Parakeet, Plum-headed Parakeet, Spotted Dove, and Vernal Hanging Parakeet, which were removed from the data because either the error rates were too high, or it was not possible to obtain estimates from photos. So, we have eye size estimates from 39 species. 

#In total, we have photo-derived eye size estimates for 36 + 39 = 75 species.
```

#Section I

## Filtering the required species and obtaining the optical axial diameter measurements for each species (From Ausprey and Ritlands's dataset)

```{r}
eye_size_museum <- eye_size_museum %>%
  rename (scientific_name = species_jetz) %>% #renaming for easier analyses
  dplyr::select (scientific_name, AD) %>% #retaining only select columns
  filter (scientific_name %in% eye_size_ebird_AR$scientific_name) %>% #filtering the required species
  group_by (scientific_name) %>%
  summarize (mean_AD = mean (AD)) %>% #obtaining the mean axial diameter measurements for our species
  left_join (., species_traits, by = 'scientific_name') %>%
  dplyr::select (scientific_name, mean_AD, common_name, mass) %>%
  filter (scientific_name %in% annotations$scientific_name) %>%
  ungroup ()
```

## Conducting the linear regression between eye size estimates/measurements and body mass 

```{r}
#macaulay-estimates:
lm_ebird <- lm (log (mean_eye_size) ~ log (mass), data = eye_size_ebird_AR) #linear regression between the logs of mean eye sizes and body mass for each species  
summary (lm_ebird) #to get the results 

predicted <- fitted (lm_ebird) #gives the predicted values 
residuals <- residuals (lm_ebird) #gives the residuals, i.e. the difference between the predicted and the actual value
lm_ebird_residuals <- as.data.frame (predicted) #creating a dataframe with the predicted values
lm_ebird_residuals$ebird_residuals <- residuals #and the residual values

#museum-measurements:
lm_museum <- lm (log (mean_AD) ~ log (mass), data = eye_size_museum)
summary (lm_museum)

predicted <- fitted (lm_museum)
residuals <- residuals (lm_museum)
lm_museum_residuals <- as.data.frame (predicted)
lm_museum_residuals$museum_residuals <- residuals

lm_results <- bind_cols (lm_museum_residuals, lm_ebird_residuals, eye_size_ebird_AR) %>%
  dplyr::select (museum_residuals, ebird_residuals, scientific_name, common_name) #combining the residual values of the eye size estimates from macaulay and measurements from museum and adding the column of species to it
```

## Conducting a correlation between the residual values from the linear regressions between estimates or measurements and body mass

```{r}
cor_plot <- ggscatter(lm_results, x = 'museum_residuals', y = 'ebird_residuals', 
          add = "reg.line", conf.int = TRUE, color = 'black',
          xlab = "Relative eye size measurements using Ausprey & Stanley 2024", ylab = "Relative eye size estimates using Macaulay Library", col = 'black', shape = 20, size = 10, label = 'common_name' , font.label = c(20, 'black'), 
          font.family = "Century Gothic", repel = TRUE) +
  stat_cor (method = 'spearman', label.x = -0.20, label.y = 0.25, color ='black', size = 15, family = "Century Gothic") +
  theme_bw () +
  theme(text = element_text (family = "Century Gothic"),
  axis.text =  element_text (size = 25, family = "Century Gothic", colour = "black"),
  axis.title = element_text (size =25, family = "Century Gothic", colour = "black"))

cor_plot #to inspect the visualization

#to save the plot
ggsave(cor_plot,
  filename = "figs/cor_plot.png",  
  width    = 15,
  height   = 12,
  units    = "in",
  dpi      = 300
)
dev.off()
```

#This plot shows a highly significant positive correlation (R = 0.75, p < 0.05) between the residuals of eye size estimates from citizen-science-derived photographs and measurements from museum specimens for 36 species. Thus, we use this methodology of estimating eye sizes from photographs for all 75 species.

#Section II

## Combining both the eye size dataframes to have a single dataframe with all eye size estimates (for 75 species)

```{r}
eye_size_ebird <- bind_rows (eye_size_ebird_AR, eye_size_ebird_not_in_AR)

write.csv (eye_size_ebird, 'results/eye_size_ebird.csv')
```

## Then, we obtain the residuals of eye size estimates of all species, i.e 75 species

```{r}
lm <- lm (log (mean_eye_size) ~ log (mass), data = eye_size_ebird) 
summary (lm) 

predicted <- fitted (lm) 
residuals <- residuals (lm) 
lm_residuals <- as.data.frame (predicted) 
lm_residuals$ebird_residuals <- residuals 

lm_results <- bind_cols (lm_residuals, eye_size_ebird) %>%
  dplyr::select (ebird_residuals, scientific_name, common_name) 

write.csv (lm_results, 'results/eye_size_residuals.csv')
```

