---
editor_options: 
  chunk_output_type: console
---

# Onset times  

In this script, we will obtain the differences in times at which species vocalizes and plot it.  

## Install required libraries  

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(showtext)
library(sf)
library(raster)
library(lubridate)
library(suncalc)
library(lutz)
library(stringr)
library(ggtext)
library(ggpubr)
library(hms)
library(lubridate)
library(ggrepel)
library(patchwork)
```

## Loading the data

```{r}
annotations <- read.csv ('results/data.csv')
sites <- read.csv("data/list-of-sites.csv")
species_codes <- read.csv("data/species-annotation-codes.csv")
eye_size_residuals <- read.csv ("results/eye_size_residuals.csv")
traits <- read.csv ('data/species-trait-dat.csv') 
territoriality <- read.csv ('data/territoriality-&-sociality-data.csv')
vegetation_structure <- read.csv ('data/2020-vegetation-data.csv')
```

## Modifying the species codes data for further analyses

```{r}
species_codes <- species_codes %>%
  rename (species = species_annotation_codes) #changing the name of the column so that it can be appended with the dataframe created in the downstream analyses
```

## Cleaning the territoriality data

```{r}
territoriality <- territoriality %>% 
   dplyr::select(Species, Territory) %>%
   rename (scientific_name = Species, territoriality = Territory) %>% #renaming it for easier downstream analyses
   mutate(scientific_name = recode(scientific_name, "Pycnonotus gularis" = "Rubigula gularis", "Ictinaetus malayensis" = "Ictinaetus malaiensis", "Dendrocopos nanus" = "Yungipicus nanus", "Megalaima malabarica" = "Psilopogon malabaricus", "Rhopocichla atriceps" = "Dumetia atriceps", "Chrysocolaptes lucidus" = "Chrysocolaptes guttacristatus", "Luscinia brunnea" = "Larvivora brunnea", "Parus aplonotus" = "Machlolophus aplonotus", "Turdoides striata" = "Argya striata", "Zoothera citrina" = "Geokichla citrina", "Turdoides subrufa" = "Argya subrufa", "Celeus brachyurus" = "Micropternus brachyurus", "Muscicapa ruficauda" = "Ficedula ruficauda", "Bubo nipalensis" = "Ketupa nipalensis", "Stigmatopelia chinensis" = "Spilopelia chinensis", "Acrocephalus aedon" = "Arundinax aedon", "Cyornis pallipes" = "Cyornis pallidipes", "Megalaima viridis" = "Psilopogon viridis", "Garrulax delesserti" = "Pterorhinus delesserti", "Acritillas indica" = "Iole indica")) #to update the names of these species as per our dataset
```

## Organizing the vegetation structure data- here, we consider foliage score which ranges from 1 to 8 indicating the presence/absence of foliage at different heights from the forest floor

```{r}
vert_strat <- vegetation_structure %>%
  dplyr::select (Site_ID, Foliage_score) %>% #retaining only selected columns
  rename (site_id = Site_ID) %>% #to rename the column
  mutate(across(everything(), ~str_replace_all(., "_", ""))) %>%
  group_by (site_id) %>%
  mutate (Foliage_score = as.numeric (Foliage_score)) %>% 
  summarise(vert_strat = mean(Foliage_score)) #this gives mean number of forest layers
```

## Extracting the nautical times for each site and date

```{r}
annotations <- annotations %>%
  filter (!species == 'UK') %>% # removing all unidentified annotations
  rename (site_id = site_code) %>% #changing the name of the column as per the site data
  left_join (., sites, by = 'site_id') %>% # joining the two dataframes
  dplyr::select(!c(notes, forest_name, Year.Active, No..saplings.planted, No..species.planted, Planted.area..ha., Distance.to.continuous.forest..m., Remnant.area..ha., Notes,)) %>% # removing unwanted columns
  rename (lat = latitude, lon = longitude) %>% # renaming the columns
  ungroup ()

annotations$date <- dmy (annotations$date) #reading the dates in dmy format

nautical_dawn <- getSunlightTimes(data = annotations, 
  keep = c(
    "sunrise", 
    "nauticalDawn"), #keeping sunrise and nautical dawn times
  tz = "Asia/Kolkata"
) %>% 
  distinct(.) 

nautical_dawn$nauticalDawn <- as_hms(nautical_dawn$nauticalDawn) #reading the times in hms format

annotations <- annotations %>%
  left_join (., nautical_dawn, by = c('lat','lon', 'date')) #now, we have a dataframe with species, their onset times, dates, nautical dawn times, and sunrise times corresponding to each annotation at each site for both the seasons. 
```

## Filtering species with adequate observations

```{r}
sp <- annotations %>%
  group_by (species) %>%
  summarize (n= n()) %>%
  filter (!n<10) %>%
  pull (species) #this helps us only retain those species for which we have atleast 10 observations, i.e. 89 species.
```

## Creating a summarized dataframe of median relative onset times with quartile ranges, and adding species-specific information to this data

```{r}
consolidated_data <- annotations %>%
  filter (species %in% sp) %>% #filtering the selected species
  dplyr::select (site_id, filename, date, begin_clock_time, species, nauticalDawn, sunrise, peak_freq) %>%

#adding data on site-wise vertifical stratification 
  left_join (., vert_strat, by = 'site_id') %>%
  
 mutate (begin_clock_time_hms = as.POSIXct(begin_clock_time, format = "%H:%M:%S")) %>% #converting to a Posixct object 
  mutate (begin_clock_time_hms = as_hms (begin_clock_time_hms)) %>% #retaining only hours and mins information

#calculating relative onset times as the differences between timing of first vocalization of each species and the nautical dawn at corresponding site, and converting it into minutes by dividing the amount in seconds by 60, before summarizing it as a median
#also, calculating the median peak frequencies of species and mean vertical stratification of all sites where the species was observed in
  mutate(relative_onset_times = as.numeric (begin_clock_time_hms-nauticalDawn), units = 'seconds', relative_onset_times = relative_onset_times / 60) %>% 
  group_by (species) %>%
  summarise (median_begin_time = median(begin_clock_time_hms), median_relative_onset_time = median(relative_onset_times), q1 = quantile(relative_onset_times, 0.25), q3 = quantile(relative_onset_times, 0.75), median_pf = median(peak_freq), mean_vert_strat = mean (vert_strat)) %>% #calculating the median onset times, median relative onset times, quartile ranges, and median peak frequencies

#joining the species annotation codes dataset
  inner_join (., species_codes, by = 'species') %>% 
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) %>% #updating the scientific name of yellow-browed bulbul

#adding relative eye size estimates/residuals of the species
  left_join (., eye_size_residuals, by = c('scientific_name', 'common_name')) %>% 
  na.omit () %>% #removing those species for which we couldn't obtain eye size data
  filter (!common_name %in% c("Red-wattled Lapwing", "Spotted Dove", "Malabar Parakeet", "Vernal Hanging-Parrot", "Plum-headed Parakeet", "Malabar Gray Hornbill")) %>% #this gives us 76 species
  mutate(scientific_name = recode(scientific_name,"Acritillas indica" =  "Iole indica")) %>% #changing the scientific name again 
  
#adding territoriality information of the species
  left_join (., territoriality, by = 'scientific_name') %>% #joining it with the onset times, i.e. outcome dataframe
  mutate(territoriality = case_when(territoriality %in% "1" ~ "Non-territorial", territoriality %in% "2" ~ "Weakly territorial", territoriality %in% "3" ~ "Highly territorial")) %>%
  
#adding trait information of foraging height, body mass, and trophic niche of the species
  left_join (., traits, by = c('scientific_name', 'common_name')) %>%
  dplyr::select (species, median_begin_time, median_relative_onset_time, scientific_name, common_name, Order, ebird_residuals, q1, q3, trophic_niche, mass,foraging_habit, median_pf, territoriality, mean_vert_strat) %>% #only keeping columns which are required for further analyses

  filter(trophic_niche != "Aquatic predator") %>%
  filter(trophic_niche != "Granivore") %>%
  filter(trophic_niche != "Nectarivore") %>%
  filter(trophic_niche != "Vertivore") %>%
  filter (foraging_habit != "AER") %>%
  filter (foraging_habit != "AQU") #removing trophic niches and foraging habit categories that are poorly represented (Aquatic predator- 1 sp; Granivore- 1 sp; Nectarivore- 2 sp; Vertivore- 3 sp; AQU- 1 sp; AER- 1 sp)
  
#This is a consolidated dataframe for 67 species which contains all the information for downstream analyses.

write.csv (consolidated_data, 'results/consolidated_data.csv')
```

## Using this consolidated data to add colours and icons to the plot

```{r}
consolidated_data <- consolidated_data %>%

#giving different colours to indicate the differences in relative onset times
  mutate (colour = case_when (median_relative_onset_time > 0 & median_relative_onset_time <= 20 ~ '#FFB000',
                              median_relative_onset_time > 20 & median_relative_onset_time <= 40 ~ '#009E73',
                              median_relative_onset_time > 40 & median_relative_onset_time <= 60 ~ '#000000',
                              median_relative_onset_time > 60 & median_relative_onset_time <= 80 ~ '#FE6100',
                              median_relative_onset_time > 80 ~ '#785EF0')) %>%
   arrange(median_relative_onset_time) %>%

#using free icons from FontAwesome to represent trophic niches
  mutate (tn_icon = case_when (trophic_niche == 'Invertivore' ~ '&#xf717;',
                               trophic_niche == 'Vertivore' ~ '&#xf4ba;',
                               trophic_niche == 'Nectarivore' ~ '&#xf5d2;',
                               trophic_niche == 'Granivore' ~ '&#xe2cd;',
                               trophic_niche == 'Aquatic predator' ~ '&#xf578;',
                               trophic_niche == 'Frugivore' ~ '&#xf5d1;',
                               trophic_niche == 'Omnivore' ~ '&#xf06c;')) %>%
  mutate (font_solid = "fa-solid") %>%
  mutate(tn_label = paste0("<span style='font-family:", font_solid, "'>", tn_icon, "</span>")) %>%
  
#using free icons from FontAwesome to represent foraging habits
  mutate (fh_icon = case_when (foraging_habit == 'CAN' ~ '&#xf111;',
                               foraging_habit == 'GRD' ~ '&#xf0c8;',
                               foraging_habit == 'MID' ~ '&#xf005;',
                               foraging_habit == 'UND' ~ '&#xe80a;',
                               foraging_habit == 'AQU' ~ '&#xf219;',
                               foraging_habit == 'AER' ~ '&#xe139;')) %>%
  mutate (fh_label = paste0("<span style='font-family:", font_solid, "'>", fh_icon, "</span>")) %>%

#using free icons from FontAwesome to represent territoriality
  mutate (te_icon = case_when (territoriality == 'Highly territorial' ~ '&#xe510;',
                               territoriality == 'Weakly territorial' ~ '&#xe50a;',
                               territoriality == 'Non-territorial' ~ '&#xe50b;')) %>%
  mutate (te_label = paste0("<span style='font-family:", font_solid, "'>", te_icon, "</span>")) %>%
  mutate(common_name_styled = paste0("<span style='color:", colour, "'>", common_name, "</span>"), #creating the HTML-styled labels for coloring
    common_name_ordered = fct_reorder(common_name, median_relative_onset_time)) #create an ordering factor
```

## Plotting all the relative onset times of 76 species to look at the overall distribution of relative onset times- here, we plainly represent the raw data, i.e. carry out the same steps as above, but without summarizing it

```{r}
#Organizing the data-
annotations <- annotations %>%
  filter (species %in% sp) %>% 
  dplyr::select (site_id, filename, date, begin_clock_time, species, nauticalDawn, sunrise) %>%
  mutate (begin_clock_time_hms = as.POSIXct(begin_clock_time, format = "%H:%M:%S")) %>% 
  mutate (begin_clock_time_hms = as_hms (begin_clock_time_hms)) %>%
  mutate(relative_onset_times = as.numeric (begin_clock_time_hms-nauticalDawn), relative_onset_times = relative_onset_times / 60) %>%
  inner_join (., species_codes, by = 'species') %>%

#adding all the species-specific information from the consolidated data
  left_join (., consolidated_data, by = c('scientific_name', 'common_name', 'Order', 'species')) %>%
  na.omit () %>%
  mutate (common_name_ordered = fct_reorder(common_name, relative_onset_times)) 


#Plotting it as a boxplot-
rel_onset_times_boxplot <- ggplot(annotations, aes(x = common_name_ordered, y = relative_onset_times, color = colour)) +
  geom_boxplot (outlier.shape = NA) +
  geom_jitter (shape = 20, size = 1.2, alpha = 0.5) +
  coord_flip() +
  theme_bw() +
  scale_x_discrete(labels = setNames(annotations$common_name_styled, annotations$common_name_ordered)) +
  scale_color_identity(guide = "legend", "Difference between onset times and nautical dawn times", labels = c('between 40 to 60 mins', 'between 20 to 40 mins', 'more than 80 mins', 'between 60 to 80 mins', 'between 0 to 20 mins')) + 
  labs(y = 'Relative onset times (minutes)', x = 'Species common name') +
  theme(
    axis.text.y = element_markdown(family = 'Century Gothic', size = 8),
    axis.text.x = element_text(family = 'Century Gothic', size = 8), 
    text = element_text(family = 'Century Gothic'),
    axis.title = element_text(family = 'Century Gothic', size = 14, face = 'bold')
  )

# inspect the visualization
rel_onset_times_boxplot

# save the plot
ggsave (rel_onset_times_boxplot,
  filename = "figs/rel_onset_times.png",  width    = 14,
  height   = 16,
  units    = "in",
  dpi      = 300
)
dev.off()
```

