---
title: "03_dag"
output: html_document
---

# DAG

In this script, we will create a directed acyclic graph based on assumptions/theories and then test the DAG-data consistency.

## Install required libraries

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(dagitty)
library(ggdag)
library(visreg)
library(patchwork)
library(ggplot2)
library(extrafont)
library(showtext)
library(scico)
library(data.table)
library(sf)
library(raster)
library(lubridate)
library(suncalc)
library(lutz)
library(ggtext)
library(ggpubr)
library(hms)
library(psych)
```

## Reading the required data

```{r}
consolidated_data <- read.csv ('results/consolidated_data.csv')
```

## Creating an acyclic graph of the conceptual model

```{r}
dag <- dagitty("
dag {
  Foraging_Height -> Eye_Size
  Body_Mass -> Eye_Size
  Vegetation_Structure -> Foraging_Height
  Vegetation_Structure -> Eye_Size
  Trophic_Niche -> Eye_Size
  Body_Mass -> Peak_Frequency
  Peak_Frequency -> Relative_onset_times
  Territoriality -> Relative_onset_times
  Eye_Size -> Relative_onset_times
  Foraging_Height -> Relative_onset_times
  Vegetation_Structure -> Peak_Frequency
  Vegetation_Structure -> Trophic_Niche
}
") #defining the dag

tidy_dag <- tidy_dagitty (dag) %>% #converting it into a tidyverse object
  mutate (type = case_when (name == 'Eye_Size' ~ 'Collider',
                            name == 'Relative_onset_times' ~ 'Response',
                            name == 'Peak_Frequency' ~ 'Collider',
                            name == 'Ambient_temperature' ~ 'Unmodelled Variable',
                            name == 'Precipitation' ~ 'Unmodelled Variable',
                            name == 'Cloud_cover' ~ 'Unmodelled Variable', .default = 'Non-collider')) #labelling the response, collider, and non-collider variables

set.seed (123)
dag_plot <- ggplot(tidy_dag, aes(x = x, y = y, xend = xend, yend = yend, color = type)) +
  geom_dag_point() +
  geom_dag_edges() +
  theme_dag () +
  geom_dag_text(family = "Century Gothic", color = 'black', size = 6, fontface = 1) +
  scale_colour_manual("Type of variable", values = c("Collider" = "#FFB000", "Non-collider" = "#009E73", "Response" = '#785EF0', "Unmodelled Variable" = "#88CCEE")) +
  theme(legend.position = 'right', 
        legend.text = element_text (size = 15, color = 'black', family = "Century Gothic"),
        legend.title = element_text (size = 15, color = 'black', family = "Century Gothic"))

dag_plot #inspecting the visualization

# save the plot
ggsave (dag_plot,
  filename = "figs/dag_plot.png",  width = 15,
  height   = 12,
  units    = "in",
  dpi      = 300, bg = "white"
)
dev.off()
```  

## Obtaining the adjustment sets to test each predictor variable with the outcome variable

```{r}
adjustmentSets(dag, exposure = "Trophic_Niche", outcome = "Relative_onset_times") #Vegetation Structure
adjustmentSets(dag, exposure = "Territoriality", outcome = "Relative_onset_times") #none
adjustmentSets(dag, exposure = "Foraging_Height", outcome = "Relative_onset_times") #Vegetation Structure
adjustmentSets(dag, exposure = "Eye_Size", outcome = "Relative_onset_times") #Foraging Height, Peak frequency; Body mass, Foraging Height, Vegetation Structure
adjustmentSets(dag, exposure = "Vegetation_Structure", outcome = "Relative_onset_times") #none
adjustmentSets(dag, exposure = "Peak_Frequency", outcome = "Relative_onset_times") #Eye size, Foraging Height; Body mass, Vegetation Structure
```

#Testing the DAG-data consistency: 

## Converting the dataframe as per testable format

```{r}
consolidated_data <- consolidated_data %>%
  dplyr::select (median_relative_onset_time, ebird_residuals, trophic_niche, foraging_habit, median_pf, territoriality, mean_vert_strat, mass, scientific_name, common_name) %>% #retaining required columns
  rename ('Eye_Size' = 'ebird_residuals', 'Relative_onset_times' = 'median_relative_onset_time', 'Trophic_Niche' = 'trophic_niche', 'Foraging_Height' = 'foraging_habit', 'Peak_Frequency' = 'median_pf', 'Territoriality' = 'territoriality', 'Vegetation_Structure' = 'mean_vert_strat', 'Body_Mass' = 'mass') #changing the names of the columns as per the DAG
```
  
## Checking DAG-data consistency  

```{r}
test <- localTests(dag, consolidated_data, type = "cis.chisq") #using chi-square test because our data has a combination of continuous and categorical variables 

test$p.value <- p.adjust (test$p.value) #to account for false-positive results from the multiple testing problem

#In this, we test if the conditional independencies implied by the model are true, in which case we are trying to see if the null hypotheses are correct. So, we are looking for non-significant p-values with effect sizes closer to 0.

plotLocalTestResults (test) #this plot shows that the relationships that we have specified in the model have effect sizes that fall on or near 0 on the line, except for three relationships involving trophic niche, peak frequency, foraging height, and vegetation structure, which show effect sizes of 1 with broad confidence intervals.
```










