---
title: "03_dag"
output: html_document
---

# DAG

In this script, we will create a directed acyclic graph based on assumptions/theories and then test the DAG-data consistency.

## Install required libraries

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(dagitty)
library(ggdag)
library(visreg)
library(patchwork)
library(ggplot2)
library(extrafont)
library(showtext)
library(scico)
library(data.table)
library(sf)
library(raster)
library(lubridate)
library(suncalc)
library(lutz)
library(ggtext)
library(ggpubr)
library(hms)
library(psych)
```

## Reading the required data

```{r}
annotations <- read.csv ('results/data.csv') 
sites <- read.csv("data/list-of-sites.csv")
summer_rel_onset_times <- read.csv ('results/summer_rel_onset_times.csv')
winter_rel_onset_times <- read.csv ('results/winter_rel_onset_times.csv')
territoriality <- read.csv ('data/territoriality-&-sociality-data.csv')
eye_size_residuals <- read.csv ('results/eye_size_residuals.csv')
traits <- read.csv ('data/species-trait-dat.csv') 
vegetation_structure <- read.csv ('data/2020-vegetation-data.csv')
species_codes <- read.csv("data/species-annotation-codes.csv")
```

#Creating and exploring the DAG:

## Creating an acyclic graph of the conceptual model

```{r}
dag <- dagitty("
dag {
  Foraging_Height -> Eye_Size
  Body_Mass -> Eye_Size
  Vegetation_Structure -> Foraging_Height
  Vegetation_Structure -> Eye_Size
  Trophic_Niche -> Eye_Size
  Body_Mass -> Peak_Frequency
  Peak_Frequency -> Relative_onset_times
  Territoriality -> Relative_onset_times
  Eye_Size -> Relative_onset_times
  Foraging_Height -> Relative_onset_times
  Vegetation_Structure -> Peak_Frequency
  Relative_onset_times [Outcome]
}
") #defining the dag

tidy_dag <- tidy_dagitty (dag) %>% #converting it into a tidyverse object
  mutate (type = case_when (name == 'Eye_Size' ~ 'Collider',
                            name == 'Relative_onset_times' ~ 'Response', .default = 'Non-collider')) #labelling the response, collider, and non-collider variables

dag_plot <- ggplot(tidy_dag, aes(x = x, y = y, xend = xend, yend = yend, color = type)) +
  geom_dag_point() +
  geom_dag_edges() +
  theme_dag () +
  geom_dag_text(family = "Century Gothic", color = 'black', size = 6, fontface = 1) +
  scale_colour_manual(values = c("Collider" = "#FFB000", "Non-collider" = "#009E73", "Response" = '#785EF0')) +
  theme(legend.position = 'none') 

# save the plot
ggsave (dag_plot,
  filename = "figs/dag_plot.png",  width = 15,
  height   = 12,
  units    = "in",
  dpi      = 300, bg = "white"
)
dev.off()
```  

##Obtaining the adjustment sets to test each predictor variable with the outcome variable

```{r}
adjustmentSets(dag, exposure = "trophic_niche", outcome = "onset_times_wrt_light_or_time") #eye size
adjustmentSets(dag, exposure = "territoriality", outcome = "onset_times_wrt_light_or_time") #none
adjustmentSets(dag, exposure = "foraging_height", outcome = "onset_times_wrt_light_or_time") #eye size, vegetation structure
adjustmentSets(dag, exposure = "eye_size", outcome = "onset_times_wrt_light_or_time") #body mass
adjustmentSets(dag, exposure = "vegetation_structure", outcome = "onset_times_wrt_light_or_time") #eye size
adjustmentSets(dag, exposure = "peak_frequency", outcome = "onset_times_wrt_light_or_time") #body mass, vegetation structure
adjustmentSets(dag, exposure = "body_mass", outcome = "onset_times_wrt_light_or_time") #none
```

#Obtaining the DAG-data consistency: 

#Outcome variable: Relative onset times

## Cleaning the onset times dataset 

```{r}
#summer
summer_rel_onset_times <- summer_rel_onset_times %>%
  dplyr::select (species, median_relative_onset_time, scientific_name, common_name) %>%
  rename ('onset_times_wrt_light_or_time' = median_relative_onset_time) #renaming the column so that it corresponds to the dag

#winter
winter_rel_onset_times <- winter_rel_onset_times %>%
  dplyr::select (species, median_relative_onset_time, scientific_name, common_name) %>%
  rename ('onset_times_wrt_light_or_time' = median_relative_onset_time)
```

#Predictor variable 1: Vegetation structure- Canopy cover:

## Cleaning the vegetation structure data and obtaining the mean of selected vegetation structure variables

```{r}
#canopy cover
canopy_cover <- vegetation_structure %>%
  dplyr::select (Site_ID, Canopy_cover) %>% #selecting only required columns
  rename (site_id = Site_ID) %>% #renaming the column
  mutate(across(everything(), ~str_replace_all(., "_", ""))) %>%
  group_by (site_id) %>%
  mutate (Canopy_cover = as.numeric (Canopy_cover)) %>% #gives multiple values of canopy cover for each site_id
  summarise(canopy_cover = mean(Canopy_cover)) #calculating mean canopy cover for each site_id

#vertical stratification
vert_strat <- vegetation_structure %>%
  dplyr::select (Site_ID, Foliage_score) %>% 
  rename (site_id = Site_ID) %>% 
  mutate(across(everything(), ~str_replace_all(., "_", ""))) %>%
  group_by (site_id) %>%
  mutate (Foliage_score = as.numeric (Foliage_score)) %>% 
  summarise(vert_strat = mean(Foliage_score))

#number of trees and unique species
treerich <- vegetation_structure %>% 
  rename (site_id = Site_ID) %>%
  mutate(across(everything(), ~str_replace_all(., "_", ""))) %>%
  group_by(site_id) %>% 
  summarise (count = n(), richness = n_distinct(tree_species))

all_Veg <- canopy_cover %>%
  left_join (treerich) %>% 
  left_join (vert_strat)

pairs.panels(all_Veg[,2:5]) #to check for correlation- the variables are correlated

#conducting PCA 
veg_pca <- prcomp(all_Veg[, 2:5], scale=TRUE, center = TRUE, retx=TRUE)
summary(veg_pca) #PC1 explains ~73% of the variance

veg_pc1 <- data.frame('site_id'= all_Veg$site_id, 'veg_pc1' = veg_pca$x[,1]) #extracting PC1 values
```

## Cleaning and organizing the onset times data 

```{r}
#summer
summer_annotations <- subset (annotations, annotations$season == 'summer') #separating out the seaons

summer_annotations <- summer_annotations %>%
  filter (!species == 'UK') %>% #to remove unidentified annotations
  rename (site_id = site_code) %>% #changing the name of the column
  left_join (., sites, by = 'site_id') %>% #adding the site information
  dplyr::select(!c(notes, forest_name, Year.Active, No..saplings.planted, No..species.planted, Planted.area..ha., Distance.to.continuous.forest..m., Remnant.area..ha., Notes,)) %>% #retaining only required columns
  rename (lat = latitude, lon = longitude) %>% #renaming the lat and long columns
  ungroup ()

summer_annotations$date <- dmy (summer_annotations$date) #so that the date column is read in dmy format

summer_nautical_dawn <- getSunlightTimes(data = summer_annotations, 
  keep = c(
    "sunrise", 
    "nauticalDawn"), 
  tz = "Asia/Kolkata"
) %>% 
  distinct(.) #to obtain sunrise and nautical dawn times

summer_nautical_dawn$nauticalDawn <- as_hms(summer_nautical_dawn$nauticalDawn) #to read in hms format 

summer_annotations <- summer_annotations %>%
  left_join (., summer_nautical_dawn, by = c('lat','lon', 'date')) #adding the nautical dawn information to the annotations data

summer_species <- summer_annotations %>%
  group_by (species) %>%
  summarize (n= n()) %>%
  filter (!n<10) %>%
  pull (species) #retaining species for which we have atleast 10 observations, i.e 84 species

#winter
winter_annotations <- subset (annotations, annotations$season == 'winter') 

winter_annotations <- winter_annotations %>%
  filter (!species == 'UK') %>% 
  rename (site_id = site_code) %>% 
  left_join (., sites, by = 'site_id') %>% 
  dplyr::select(!c(notes, forest_name, Year.Active, No..saplings.planted, No..species.planted, Planted.area..ha., Distance.to.continuous.forest..m., Remnant.area..ha., Notes,)) %>% 
  rename (lat = latitude, lon = longitude) %>% 
  ungroup ()

winter_annotations$date <- dmy (winter_annotations$date) 

winter_nautical_dawn <- getSunlightTimes(data = winter_annotations, 
  keep = c(
    "sunrise", 
    "nauticalDawn"), 
  tz = "Asia/Kolkata"
) %>% 
  distinct(.) 

winter_nautical_dawn$nauticalDawn <- as_hms(winter_nautical_dawn$nauticalDawn) #to read in hms format 

winter_annotations <- winter_annotations %>%
  left_join (., winter_nautical_dawn, by = c('lat','lon', 'date')) 

winter_species <- winter_annotations %>%
  group_by (species) %>%
  summarize (n= n()) %>%
  filter (!n<10) %>%
  pull (species) #gives 74 species
```

## Adding the vegetation structure data to the onset times data

```{r}
species_codes <- species_codes %>%
  rename (species = species_annotation_codes) #to add scientific names and common names to the dataframe

#summer
summer_vegetation_structure <- summer_annotations %>%
  filter (species %in% summer_species) %>% #filtering those species
  dplyr::select (season, site_id, filename, date, begin_clock_time, species, nauticalDawn, sunrise) %>% #retaining only required columns
  left_join (., veg_pc1, by = 'site_id') %>% #adding vegetation structure pc1 data
  mutate (begin_clock_time_hms = as.POSIXct(begin_clock_time, format = "%H:%M:%S")) %>% 
  mutate (begin_clock_time_hms = as_hms (begin_clock_time_hms)) %>%
  mutate(relative_onset_times = as.numeric (begin_clock_time_hms - nauticalDawn), units = 'seconds') %>%   group_by (species) %>%
  summarise (median_begin_time = median(begin_clock_time_hms), median_relative_onset_time = median(relative_onset_times),vegetation_structure = mean (veg_pc1)) %>% #calculating relative onset times as the differences between timing of first vocalization of each species and the nautical dawn at corresponding site
  left_join (., species_codes, by = 'species') %>% #adding the columns of scientific name and common name
  dplyr::select (species, scientific_name, vegetation_structure, common_name) %>% #retaining only required columns
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) #updating the scientific name of yellow-browed bulbul

#winter
winter_vegetation_structure <- winter_annotations %>%
  filter (species %in% winter_species) %>% 
  dplyr::select (season, site_id, filename, date, begin_clock_time, species, nauticalDawn, sunrise) %>% 
  left_join (., veg_pc1, by = 'site_id') %>% 
  mutate (begin_clock_time_hms = as.POSIXct(begin_clock_time, format = "%H:%M:%S")) %>% 
  mutate (begin_clock_time_hms = as_hms (begin_clock_time_hms)) %>%
  mutate(relative_onset_times = as.numeric (begin_clock_time_hms - nauticalDawn), units = 'seconds') %>%   group_by (species) %>%
  summarise (median_begin_time = median(begin_clock_time_hms), median_relative_onset_time = median(relative_onset_times), vegetation_structure = mean (veg_pc1)) %>% 
  left_join (., species_codes, by = 'species') %>% 
  dplyr::select (species, scientific_name, vegetation_structure, common_name) %>% 
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) 

#This gives us dataframes with species and the mean PC1 scores of vegetation structure variables of the sites where the species vocalized in, for each season.
```

#Predictor variable 2: Peak Frequency:

## Obtaining the median peak frequencies of species

```{r}
#summer
summer_peak_freq <- annotations %>%
  group_by (species) %>%
  summarise(median_peak_freq = median(peak_freq)) %>% #obtaining the median peak frequency and
  mutate (peak_frequency = log10(median_peak_freq)) %>% #taking a log of it
  left_join (summer_rel_onset_times, ., by = 'species' ) %>% #joining it with the onset times, i.e. outcome dataframe
  dplyr::select (peak_frequency, species, scientific_name, common_name) #retaining only required columns

#winter
winter_peak_freq <- annotations %>%
  group_by (species) %>%
  summarise(median_peak_freq = median(peak_freq)) %>% 
  mutate (peak_frequency = log10(median_peak_freq)) %>% 
  left_join (winter_rel_onset_times, ., by = 'species' ) %>% 
  dplyr::select (peak_frequency, species, scientific_name, common_name) 

#This gives us dataframes with species and their log median peak frequencies for each season.
```

#Predictor variable 3: Trophic niche:

## Obtaining the trophic niche of species

```{r}
#summer
summer_trophic_niche <- traits %>%
  dplyr::select (scientific_name, common_name, trophic_niche) %>% #selecting only required columns
  left_join (summer_rel_onset_times, ., by = c('common_name', 'scientific_name')) %>% #joining it with the onset times, i.e. outcome dataframe
  rename ('trophic_niche' = trophic_niche) %>% #renaming the column so that it corresponds to the dag
  dplyr::select (trophic_niche, species, scientific_name, common_name) %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) #updating the scientific name of yellow-browed bulbul

#winter
winter_trophic_niche <- traits %>%
  dplyr::select (scientific_name, common_name, trophic_niche) %>% 
  left_join (winter_rel_onset_times, ., by = c('common_name', 'scientific_name')) %>% 
  rename ('trophic_niche' = trophic_niche) %>% 
  dplyr::select (trophic_niche, species, scientific_name, common_name) %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) 

#This gives us dataframes with species and their trophic niches for each season.
```

#Predictor variable 3: Foraging Height:

## Obtaining the foraging height of species

```{r}
#summer
summer_foraging_height <- traits %>%
  dplyr::select (scientific_name, common_name, foraging_habit) %>% #selecting only required columns
  left_join (summer_rel_onset_times, ., by = c('common_name', 'scientific_name')) %>% #joining it with the onset times, i.e. outcome dataframe
  rename ('foraging_height' = foraging_habit) %>% #renaming the column so that it corresponds to the dag
  dplyr::select (foraging_height, species, scientific_name, common_name) %>% 
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) #updating the scientific name of yellow-browed bulbul

#winter
winter_foraging_height <- traits %>%
  dplyr::select (scientific_name, common_name, foraging_habit) %>% 
  left_join (winter_rel_onset_times, ., by = c('common_name', 'scientific_name')) %>% 
  rename ('foraging_height' = foraging_habit) %>% 
  dplyr::select (foraging_height, species, scientific_name, common_name) %>% 
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica"))

#This gives us dataframes with species and their foraging heights for each season.
```

#Predictor variable 4: Territoriality:

## Obtaining territoriality data of species

```{r}
territoriality <- territoriality %>% 
   dplyr::select(Species, Territory) %>%
   rename (scientific_name = Species, territoriality = Territory) %>% #renaming it for easier downstream analyses
   mutate(scientific_name = recode(scientific_name, "Pycnonotus gularis" = "Rubigula gularis", "Ictinaetus malayensis" = "Ictinaetus malaiensis", "Dendrocopos nanus" = "Yungipicus nanus", "Megalaima malabarica" = "Psilopogon malabaricus", "Rhopocichla atriceps" = "Dumetia atriceps", "Chrysocolaptes lucidus" = "Chrysocolaptes guttacristatus", "Luscinia brunnea" = "Larvivora brunnea", "Parus aplonotus" = "Machlolophus aplonotus", "Turdoides striata" = "Argya striata", "Zoothera citrina" = "Geokichla citrina", "Turdoides subrufa" = "Argya subrufa", "Celeus brachyurus" = "Micropternus brachyurus", "Muscicapa ruficauda" = "Ficedula ruficauda", "Bubo nipalensis" = "Ketupa nipalensis", "Stigmatopelia chinensis" = "Spilopelia chinensis", "Acrocephalus aedon" = "Arundinax aedon", "Cyornis pallipes" = "Cyornis pallidipes", "Megalaima viridis" = "Psilopogon viridis", "Garrulax delesserti" = "Pterorhinus delesserti", "Acritillas indica" = "Iole indica")) #to update the names of these species as per our dataset

#summer
summer_territoriality <-  left_join (summer_rel_onset_times, territoriality, by = 'scientific_name') %>% #joining it with the onset times, i.e. outcome dataframe
  mutate(Territoriality = case_when(territoriality %in% "1" ~ "Non-territorial", territoriality %in% "2" ~ "Weakly territorial", territoriality %in% "3" ~ "Highly territorial")) %>%
  dplyr::select (territoriality, species, scientific_name, common_name) %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) #updating the scientific name of yellow-browed bulbul

#winter
winter_territoriality <-  left_join (winter_rel_onset_times, territoriality, by = 'scientific_name') %>%
  mutate(Territoriality = case_when(territoriality %in% "1" ~ "Non-territorial", territoriality %in% "2" ~ "Weakly territorial", territoriality %in% "3" ~ "Highly territorial")) %>%
  dplyr::select (territoriality, species, scientific_name, common_name) %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica"))

#These dataframes give us the species and their territorialities for each season. 
```

#Predictor variable 5: Body mass:

## Obtaining the body mass of species

```{r}
#summer
summer_body_mass <- traits %>%
  left_join (summer_rel_onset_times, ., by = c("scientific_name", "common_name")) %>%
  dplyr::select (common_name, scientific_name, mass) %>% #only retaining required columns
  rename ('body_mass' = mass) %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) #updating the scientific name of yellow-browed bulbul

#winter
winter_body_mass <- traits %>%
  left_join (winter_rel_onset_times, ., by = c("scientific_name", "common_name")) %>%
  dplyr::select (common_name, scientific_name, mass) %>% 
  rename ('body_mass' = mass) %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) 
  
#These dataframes give us the species and their body masses for each season. 
```

#Predictor variable 6: Eye Size:

## Adding eye size for each species

```{r}
#summer
summer_eye_size <- eye_size_residuals %>%
  left_join (., summer_rel_onset_times, by = c('scientific_name', 'common_name')) %>% #joining it with the onset times, i.e. outcome dataframe
  dplyr::select (species, common_name, scientific_name, onset_times_wrt_light_or_time, ebird_residuals) %>% #retaining only required columns
  rename ('eye_size' = ebird_residuals) %>%
  dplyr::select (eye_size, species, common_name, scientific_name)

#winter
winter_eye_size <- eye_size_ebird %>%
  left_join (., winter_rel_onset_times, by = c('scientific_name', 'common_name')) %>% 
  dplyr::select (species, common_name, scientific_name, onset_times_wrt_light_or_time, rel_mean_eye_size) %>% 
  rename ('eye_size' = rel_mean_eye_size) %>%
  dplyr::select (eye_size, species, common_name, scientific_name)

#These dataframes give the species and the mean eye sizes relative to their body masses for each season.
```

## Combining all the predictor and outcome variables to a single dataframe

```{r}
#summer
summer_rel_onset_times <- summer_rel_onset_times %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) #updating the scientific name of yellow-browed bulbul

#combining all dataframes  
summer_dfs <- list (summer_rel_onset_times, summer_territoriality, summer_eye_size, summer_body_mass, summer_peak_freq, summer_foraging_height, summer_trophic_niche, summer_vegetation_structure)

summer_dag_df <- purrr::reduce(summer_dfs, dplyr::left_join, by = c("scientific_name", "common_name")) %>%
  mutate (foraging_height = as.factor (foraging_height), trophic_niche = as.factor (trophic_niche), territoriality = as.factor (territoriality)) %>% #reading categorical variables as factors
  na.omit () %>% #to remove species for which we don't have photo-derived eye size data, which are: Brown Boobook, Great Hornbill, Indian Scops Owl, Jungle Nightjar, Malabar Parakeet, Malabar Gray Hornbill, Plum-headed Parakeet, Red-wattled Lapwing, Spot-bellied Eagle Owl, Spotted Dove, and Vernal Hanging-Parrot.
  dplyr::select (scientific_name, common_name, onset_times_wrt_light_or_time, foraging_height, trophic_niche, territoriality, body_mass, eye_size, vegetation_structure, peak_frequency) #retaining required columns

#winter
winter_rel_onset_times <- winter_rel_onset_times %>%
  mutate(scientific_name = recode(scientific_name, "Iole indica" = "Acritillas indica")) 

#combining all dataframes  
winter_dfs <- list (winter_rel_onset_times, winter_territoriality, winter_eye_size, winter_body_mass, winter_peak_freq, winter_foraging_height, winter_trophic_niche, winter_vegetation_structure)

winter_dag_df <- purrr::reduce(winter_dfs, dplyr::left_join, by = c("scientific_name", "common_name")) %>%
  mutate (foraging_height = as.factor (foraging_height), trophic_niche = as.factor (trophic_niche), territoriality = as.factor (territoriality)) %>% 
  na.omit () %>% #to remove species for which we don't have photo-derived eye size data, which are: Brown Boobook, Great Hornbill, Indian Scops Owl, Jungle Nightjar, Malabar Parakeet, Malabar Gray Hornbill, Plum-headed Parakeet, Red-wattled Lapwing, Spot-bellied Eagle Owl, Spotted Dove, and Vernal Hanging-Parrot.
  dplyr::select (scientific_name, common_name, onset_times_wrt_light_or_time, foraging_height, trophic_niche, territoriality, body_mass, eye_size, vegetation_structure, peak_frequency) 
```

## Evaluating the dependencies and independencies specificed in our dag

```{r}
#summer
summer_test <- localTests(dag, summer_dag_df, type = "cis.chisq") #to account for categorical data
summer_test$p.value <- p.adjust(summer_test$p.value) #this shows p-values > 0.05, ensuring dag-data consistency

plotLocalTestResults(summer_test) #to visualize the results

#winter
winter_test <- localTests(dag, winter_dag_df, type = "cis.chisq") 
winter_test$p.value <- p.adjust(winter_test$p.value) #this shows p-values > 0.05, ensuring dag-data consistency

plotLocalTestResults(winter_test)
```







