---
title: "03_dag"
output: html_document
---

# DAG

In this script, we will create a directed acyclic graph based on assumptions/theories and then test the DAG-data consistency.

## Install required libraries

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(dagitty)
library(ggdag)
library(visreg)
library(patchwork)
library(ggplot2)
library(extrafont)
library(showtext)
library(scico)
library(data.table)
library(sf)
library(raster)
library(lubridate)
library(suncalc)
library(lutz)
library(ggtext)
library(ggpubr)
library(hms)
library(psych)
```

## Reading the required data

```{r}
consolidated_data <- read.csv ('results/consolidated_data.csv')
```

## Cleaning the data

```{r}
consolidated_data <- consolidated_data %>%
  filter(trophic_niche != "Aquatic predator") %>%
  filter(trophic_niche != "Granivore") %>%
  filter(trophic_niche != "Nectarivore") %>%
  filter(trophic_niche != "Vertivore") %>%
  filter (foraging_habit != "AER") %>%
  filter (foraging_habit != "AQU") #removing trophic niches and foraging habit categories that are poorly represented (Aquatic predator- 1 sp; Granivore- 1 sp; Nectarivore- 2 sp; Vertivore- 3 sp; AQU- 1 sp; AER- 1 sp)
  
#This leaves us with 67 species for downstream analyses.
```


## Creating an acyclic graph of the conceptual model

```{r}
dag <- dagitty("
dag {
  Foraging_Height -> Eye_Size
  Vegetation_Structure -> Foraging_Height
  Vegetation_Structure -> Eye_Size
  Vegetation_Structure -> Relative_onset_times
  Trophic_Niche -> Eye_Size
  Body_Mass -> Peak_Frequency
  Trophic_Niche -> Foraging_Height
  Peak_Frequency -> Relative_onset_times
  Territoriality -> Relative_onset_times
  Eye_Size -> Relative_onset_times
  Foraging_Height -> Relative_onset_times
  Vegetation_Structure -> Peak_Frequency
  Vegetation_Structure -> Trophic_Niche
}
") #defining the dag
#Body_Mass -> Eye_Size- removed since we are using eye size residuals

tidy_dag <- tidy_dagitty (dag) #converting it into a tidyverse object

set.seed (123)
dag_plot <- ggplot(tidy_dag, aes(x = x, y = y, xend = xend, yend = yend)) +#color = type)) +
  geom_dag_point() +
  geom_dag_edges() +
  theme_dag () +
  geom_dag_text(family = "Century Gothic", color = 'red', size = 6, fontface = 1) +
  theme(legend.position = 'right', 
        legend.text = element_text (size = 15, color = 'black', family = "Century Gothic"),
        legend.title = element_text (size = 15, color = 'black', family = "Century Gothic"))

dag_plot #inspecting the visualization

# save the plot
ggsave (dag_plot,
  filename = "figs/dag_plot.png",  width = 15,
  height   = 12,
  units    = "in",
  dpi      = 300, bg = "white"
)
dev.off()
```  

## Obtaining the adjustment sets to test each predictor variable with the outcome variable

```{r}
adjustmentSets(dag, exposure = "Trophic_Niche", outcome = "Relative_onset_times", effect = "direct") #Eye_Size, Foraging_Height, Vegetation_Structure
adjustmentSets(dag, exposure = "Territoriality", outcome = "Relative_onset_times", effect = "direct") #none
adjustmentSets(dag, exposure = "Foraging_Height", outcome = "Relative_onset_times", effect = "direct") #Eye Size, Vegetation Structure
adjustmentSets(dag, exposure = "Eye_Size", outcome = "Relative_onset_times", effect = "direct") #Foraging Height, Vegetation Structure
adjustmentSets(dag, exposure = "Vegetation_Structure", outcome = "Relative_onset_times", effect = "direct") #Eye Size, Foraging Height, Peak Frequency
adjustmentSets(dag, exposure = "Peak_Frequency", outcome = "Relative_onset_times", effect = "direct") #Vegetation Structure
```

#Testing the DAG-data consistency: 

## Converting the dataframe as per testable format

```{r}
consolidated_data <- consolidated_data %>%
  dplyr::select (median_relative_onset_time, ebird_residuals, trophic_niche, foraging_habit, median_pf, territoriality, mean_vert_strat, mass, scientific_name, common_name) %>% #retaining required columns
  mutate (body_mass = log(mass), median_pf_z = scale (median_pf), mean_vert_strat_z = scale (mean_vert_strat), ebird_residuals_z = scale (ebird_residuals)) %>% #scaling all the numerical variables
  rename ('Eye_Size' = 'ebird_residuals_z', 'Relative_onset_times' = 'median_relative_onset_time', 'Trophic_Niche' = 'trophic_niche', 'Foraging_Height' = 'foraging_habit', 'Peak_Frequency' = 'median_pf_z', 'Territoriality' = 'territoriality', 'Vegetation_Structure' = 'mean_vert_strat_z', 'Body_Mass' = 'body_mass') #changing the names of the columns as per the DAG
```
  
## Checking DAG-data consistency  

```{r}
test <- localTests(dag, consolidated_data, type = "cis.chisq") #using chi-square test because our data has a combination of continuous and categorical variables 

test$p.value <- p.adjust (test$p.value) #to account for false-positive results from the multiple testing problem

#In this, we test if the conditional independencies implied by the model are true, in which case we are trying to see if the null hypotheses are correct. So, we are looking for non-significant p-values with effect sizes closer to 0.

plotLocalTestResults (test) #this plot shows conditional independence statements on the y-axis and test statistic on the x-axis. Those variables for which the points are at 0- the conditional independencies (A|B: A is independent of B; A||B|C: A is independent of B given C) specified by the dag hold true for the data as well. Those variables, specifically i) independence between Territoriality and Trophic niche, ii) Peak frequency and Trophic niche, given Vegetation structure, iii) Foraging Height and Peak Frequency, given Vegetation structure, and iv) Eye size and Peak frequency, given Vegetation structure, suggests that the conditional independence specified in the dag doesn't hold true for the data. Territoriality and Trophic niche show a narrow confidence interval hinting at dag-data inconsistency, whereas the others show wide confidence intervals hints at incconclusive evidence at dag-data inconsistency.
```










